---
title: "HW 02"
subtitle: "Due Thursday, October 15, 11:59 PM"
author: "Matthew Xiao"
output: pdf_document
---

```{r load-packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
licorice <- read_csv("data/licorice.csv") 
```

### Exercise 1

```{r ex-1}
c_licorice <- licorice %>% filter(!is.na(pacu30min_throatPain))

n_sims <- 1000
set.seed(1)
boot_dist = numeric(n_sims)

for (i in 1:n_sims){
  set.seed(i)
  indices <- sample(1:nrow(c_licorice), replace = TRUE)
  boot_mean <- c_licorice %>%
    slice(indices) %>%
    summarize(boot_meean = mean(pacu30min_throatPain)) %>% pull()
  boot_dist[i] = boot_mean
  
  
  
}
boot_means <- tibble(boot_dist)
boot_means %>% summarize(Lower = quantile(boot_dist, 0.025),
                         Upper = quantile(boot_dist, 0.975)) %>% 
  knitr::kable()
t.test(c_licorice$pacu30min_throatPain, conf.level = .95)
```
Both the clt- based apprach and the simulation approach created similar bounds 
of around .49 for the lower bound and .81 for the upper bounds. 
The bounds for the simulation approach are slightly smaller as the lower 
bound is slightly higher as well as the upper bound is as well. In both cases 
the 95% constructed region creates bounds of the pain level of .487 to .809. 
If we would randomly sample the patient data for throat pain, following the same
approach we would be 95% confident that 95% of the new constructed confidence
intervals to encompass the true mean. 








### Exercise 2

```{r ex-2 base}

b_licorice <- licorice %>% summarize(mean_b = mean(preOp_calcBMI))

set.seed(2)
n_sims <- 2000
boot_dist = numeric(n_sims) 
for(i in 1:n_sims){ 
  set.seed(i) 
  indices <- sample(1:nrow(licorice), replace = T)
  boot_mean <- licorice %>%
    slice(indices) %>%                    
    summarize(boot_mean = mean(preOp_calcBMI)) %>% 
    pull()
  
  boot_dist[i] <- boot_mean 
} 
mu_0 = 26
boot_means <- tibble(boot_dist)

offset <- boot_means %>% 
  summarize(mu_0 - mean(boot_dist) )%>% pull()
boot_means <- tibble(boot_dist)
boot_means %>% summarize(Lower = quantile(boot_dist, 0.025),
                         Upper = quantile(boot_dist, 0.975)) %>% 
  knitr::kable()
boot_means <- boot_means %>% 
  mutate(shifted_means = boot_dist+offset)
ggplot(data = boot_means, aes(x = shifted_means)) +
  geom_histogram(binwidth = 0.1, color = "darkblue", fill = "skyblue") + 
  labs(x = "Mean of BMI", 
       y = "Count",
       title = "The graph displays the null distribution for 
       the mean BMI of the Preoperation Patients with the asumption that the 
       mean BMI is 26.") +
  geom_vline(xintercept = quantile(boot_means$shifted_means, 0.025), 
             lwd = 2, 
             color = "black") +
  geom_vline(xintercept = quantile(boot_means$shifted_means, 0.975),
             lwd= 2,
             color = "black")


obs <- licorice %>% 
  summarize(mean(preOp_calcBMI)) %>%
  pull()
dif <- mu_0 - obs
boot_means <- boot_means %>%
  mutate(shifted_means = boot_dist + offset)

boot_means1 <- boot_means %>%
  mutate(extreme = ifelse(shifted_means <=obs| 
                            shifted_means >=mu_0 + dif, 1,0))%>%
  count(extreme) %>% 
  mutate(prob = n / sum(n)) %>% 
  select(prob) %>% 
  slice(2)
boot_means1 %>% knitr::kable()

```

Based on the simulated data. It seems as if that the probability that the mean
of sampled patients' BMI have a 25.59 or more extreme BMI is 14.4%. This is 
above the general pvalue of .05 which means that we have insufficient evidence 
to reject the null hypothesis (that the BMI of patients is the same as the rest 
of the German population. )




### Exercise 3

```{r ex-3}

B_pain <- c_licorice %>% 
  mutate(asapain = ifelse(pacu30min_throatPain >0, 1, 0)) 


t.test(B_pain$preOp_asa ~ asapain, 
       data = B_pain,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)







```

### Exercise 4

```{r ex-4}
t.test(pacu30min_throatPain ~ treat, 
       data = c_licorice,
       mu = 0,
       var.equal = FALSE,
       alternative = "greater",
       conf.level = 0.95)


```

Those that were treated with sugar (treat 0) had a higher incidence of having 
throat pain 30 after. We recieved a pvalue of 1.8e-06 which is much smaller than 
0.05 thus we can reject the null hypothesis and state that we have evidence that
licorice decreases the incidence of having throat pain 30 min after the
operation.




### Exercise 5

```{r ex-5}
set.seed(5)
cc_licorice <- c_licorice %>% 
  mutate(pain = ifelse(pacu30min_throatPain >0 | pacu30min_swallowPain >0, 1, 0)) 
sugar <- cc_licorice %>% 
  filter(treat == 0 ) 
licorice_l <- cc_licorice %>% 
  filter(treat == 1 )
n_sims <- 2000
boot_diffs <- numeric(n_sims)


for(i in 1:n_sims){
  # create indices
  indices_h <- sample(1:nrow(sugar), replace = T)
  indices_p <- sample(1:nrow(licorice_l), replace = T)
  # bootstrap est. group means
  temp_h <- sugar %>% 
    slice(indices_h) %>% 
    summarize(mean1 = mean(pain)) %>% 
    select(mean1) %>% 
    pull()
  temp_p <- licorice_l %>% 
    slice(indices_p) %>% 
    summarize(mean_jitter = mean(pain)) %>% 
    select(mean_jitter) %>% 
    pull()

  boot_diffs[i] <- temp_h - temp_p
}
boot_diffs <- tibble(diffs = boot_diffs)
boot_diffs %>% 
  summarize(lower = quantile(diffs, 0.025),
            upper = quantile(diffs, 0.975))
obs_diff <- boot_diffs %>% 
  summarize(obs_diff = mean(diffs)) %>% 
  pull()
offset <- boot_diffs %>% 
  summarize(offset = 0 - mean(diffs)) %>% 
  pull()
null_dist <- boot_diffs %>% 
  mutate(centered_diffs = diffs + offset) %>% 
  select(centered_diffs)
null_dist %>% 
  mutate(extreme = ifelse(centered_diffs > abs(obs_diff), 1, 0)) %>% 
  summarize(p_val = mean(extreme))

```


The pvalue created is 0.0055 which is much smaller than the alpha level of 0.05.
thus we have evidence that to reject the null hypthesis and conclude that those
who take licorice in general will less likley have pain 30 minutes after. 



### Exercise 6

We do  have enough evidence. we only know that 30 minutes after that a licorice
gargle will reduce pain. However we can not tell if that hold true 90 minutes or
4 hours or at 1 am. That would require further extrapolation of the data in which 
then we can conclude with the data we collected that a licorice gargle will 
reduce the pain that a patient felt in general. 

### Exercise 7
TRUE


### Exercise 8

FALSE. Our p-value can not be evidence for, but rather it would only be evidence
that does not reject it can not support. 




### Exercise 9


True




### Exercise 10

TRUE


### Exercise 11

FALSE we can not know for sure


### Exercise 12


TRUE
