---
title: "HW 02"
subtitle: "Due Thursday, October 15, 11:59 PM"
author: "Matthew Xiao"
output: pdf_document
---

```{r load-packages-data, warning=FALSE, message=FALSE}
library(tidyverse)
licorice <- read_csv("data/licorice.csv") 
```

### Exercise 1



```{r ex-1}
c_licorice <- licorice %>% filter(!is.na(pacu30min_throatPain))

n_sims <- 1000
set.seed(1)
boot_dist = numeric(n_sims)

for (i in 1:n_sims){
  set.seed(i)
  indices <- sample(1:nrow(c_licorice), replace = TRUE)
  boot_mean <- c_licorice %>%
    slice(indices) %>%
    summarize(boot_meean = mean(pacu30min_throatPain)) %>% pull()
  boot_dist[i] = boot_mean
  
  
  
}
boot_means <- tibble(boot_dist)
boot_means %>% summarize(Lower = quantile(boot_dist, 0.025),
                         Upper = quantile(boot_dist, 0.975)) %>% 
  knitr::kable()
t.test(c_licorice$pacu30min_throatPain, conf.level = .95)
```
Both the clt- based apprach and the simulation approach created similar bounds 
of around .49 for the lower bound and .81 for the upper bounds. 
The bounds for the simulation approach are slightly smaller as the lower 
bound is slightly higher as well as the upper bound is as well. In both cases 
the 95% constructed region creates bounds of the pain level of .487 to .809. 
If we would randomly sample the patient data for throat pain, following the same
approach we would be 95% confident that 95% of the new constructed confidence
intervals to encompass the true mean. 


### Exercise 2

significance level: $\alpha$ = 0.05
$H_0$ : $\mu$ = 26
$H_A$ : $\mu$ $\neq$ 26

```{r ex-2 base}

b_licorice <- licorice %>% summarize(mean_b = mean(preOp_calcBMI))

set.seed(2)
n_sims <- 2000
boot_dist = numeric(n_sims) 
for(i in 1:n_sims){ 
  set.seed(i) 
  indices <- sample(1:nrow(licorice), replace = T)
  boot_mean <- licorice %>%
    slice(indices) %>%                    
    summarize(boot_mean = mean(preOp_calcBMI)) %>% 
    pull()
  
  boot_dist[i] <- boot_mean 
} 
mu_0 = 26
boot_means <- tibble(boot_dist)

offset <- boot_means %>% 
  summarize(mu_0 - mean(boot_dist) )%>% pull()
boot_means <- tibble(boot_dist)
boot_means %>% summarize(Lower = quantile(boot_dist, 0.025),
                         Upper = quantile(boot_dist, 0.975)) %>% 
  knitr::kable()
boot_means <- boot_means %>% 
  mutate(shifted_means = boot_dist+offset)
ggplot(data = boot_means, aes(x = shifted_means)) +
  geom_histogram(binwidth = 0.1, color = "darkblue", fill = "skyblue") + 
  labs(x = "Mean of BMI", 
       y = "Count",
       title = "The graph displays the null distribution for 
       the mean BMI of the Preoperation Patients with the asumption that the 
       mean BMI is 26") +
  geom_vline(xintercept = quantile(boot_means$shifted_means, 0.025), 
             lwd = 2, 
             color = "black") +
  geom_vline(xintercept = quantile(boot_means$shifted_means, 0.975),
             lwd= 2,
             color = "black") +
  geom_vline(xintercept = quantile(mean(boot_dist), label = "Sample BMI", 0.975),
             lwd= 2,
             color = "blue")


obs <- licorice %>% 
  summarize(mean(preOp_calcBMI)) %>%
  pull()
dif <- mu_0 - obs
boot_means <- boot_means %>%
  mutate(shifted_means = boot_dist + offset)

boot_means1 <- boot_means %>%
  mutate(extreme = ifelse(shifted_means <=obs| 
                            shifted_means >=mu_0 + dif, 1,0))%>%
  count(extreme) %>% 
  mutate(Prob = n / sum(n)) %>% 
  select(Prob) %>% 
  slice(2)
boot_means1 %>% knitr::kable()

```

The blue line is the sample BMI of the patient's. 
Based on the simulated data. It seems as if that the probability that the mean
of sampled patients' BMI have a 25.59 or more extreme BMI is 14.4%. This is 
above the general alpha level of .05 which means that we have insufficient 
evidence to reject the null hypothesis (that the BMI among patients is the same 
as the rest of the German adult population. )

We find that the p value of our test is 0.144. Since our p-value is greater than
the alpha level of 0.05, we decide to fail to reject the null hypothesis that 
the mean patient BMI is different to the mean of the general German population.

In the visualization of the graph we can see that the blue line is within the 
two black lines which are the boundaries of a 95% significance level. Thus our 
data is not extreme enough to suggest that the difference between the German BMI
of the entire population of 26 and the sample's mean is significant enough to 
reject the null hypothesis. 


We ultimately conclude that there is not enough evidence to suggest the mean 
patients have a different than the German population.



### Exercise 3
significance level:$\alpha$ = 0.05
$H_0$: Throat pain 30 minutes after and ASA classification are independent; 
  there is no association between them
$H_1$: Throat pain 30 minutes after and ASA classification are not independent; 
  there is association between them

```{r ex-3}

B_pain <- c_licorice %>% 
  mutate(asapain = ifelse(pacu30min_throatPain >0, 1, 0)) 


chisq.test(table(B_pain$asapain, B_pain$preOp_asa))

```
The distribution of the test statistic under the null distribution is a 
chi-square distribution that has a 2 degree of freedoms. 

The X^2 value of .2985. 

We find that the p value of our test is .2985 Since our p-value is greater than 
the alpha level of 0.05, we decide to fail to reject the null hypothesis that 
Throat pain 30 minutes after and ASA classification are independent. 

We ultimately conclude that there is not enough evidence to suggest that there is 
a correlation between between throat pain 30 minutes after and the ASA physical 
classification 

### Exercise 4

$\alpha$ = 0.05
$\mu_s$ : mean throat pain 30 minutes level after the operation for those with the 
sugar placebo
$\mu_l$ : mean throat pain level 30 minutes after the operation for those with 
the licorice
$H_0$: Gargling licorice does not affect the mean throat pain of patients
$H_0$ : $\mu_s$ = $\mu_l$

$H_A$: Gargling licorice decreases the the mean throat pain of patients
$H_A$ : $\mu_s$ > $\mu_l$


```{r ex-4}
t.test(pacu30min_throatPain ~ treat, 
       data = c_licorice,
       mu = 0,
       var.equal = FALSE,
       alternative = "greater",
       conf.level = 0.95)

```

Those that were treated with sugar (treat 0) had a higher incidence of having 
throat pain 30 after. We received a pvalue of 1.8e-06 which is much smaller than 
0.05 thus we can reject the null hypothesis and state that we have evidence that
licorice decreases the incidence of having throat pain 30 min after the
operation.


The distribution of the test statistic under the null distribution is a t 
distribution that has a 157.3 degrees of freedoms.

The test statistic is 4.805 

We find that the p value of our test is 1.804e-06. Since our p-value is less than
the alpha level of 0.05, we decide to reject the null hypothesis that the mean 
throat pain after 30 after a licorice gargle is greater than or equal to the  
mean throat pain after 30 with the sugar gargle, in favor of the alternative 
hypothesis that the mean throat pain after 30 after a licorice gargle is less 
than to the mean throat pain after 30 with the sugar gargle.

We ultimately conclude that there is enough evidence to suggest a licorice gargle 
decreases the mean throat pain after 30 after a licorice gargle.




### Exercise 5

significance level:$\alpha$ = 0.05
$\mu_s$ : mean pain 30 minutes probability after for those with the sugar placebo
$\mu_l$ : mean pain 30 minutes probability after for those with the licorice
$H_0$: Gargling licorice does not affect the mean throat pain of patients
$H_0$ : $\mu_s$ = $\mu_l$

$H_A$: Gargling licorice decreases the the mean throat pain of patients
$H_A$ : $\mu_s$ > $\mu_l$


```{r ex-5}
set.seed(5)
cc_licorice <- c_licorice %>% 
  mutate(pain = ifelse(pacu30min_throatPain >0 | pacu30min_swallowPain >0, 1, 0)) 
sugar <- cc_licorice %>% 
  filter(treat == 0 ) 
licorice_l <- cc_licorice %>% 
  filter(treat == 1 )
n_sims <- 2000
boot_diffs <- numeric(n_sims)


for(i in 1:n_sims){
  # create indices
  indices_h <- sample(1:nrow(sugar), replace = T)
  indices_p <- sample(1:nrow(licorice_l), replace = T)
  # bootstrap est. group means
  temp_h <- sugar %>% 
    slice(indices_h) %>% 
    summarize(mean1 = mean(pain)) %>% 
    select(mean1) %>% 
    pull()
  temp_p <- licorice_l %>% 
    slice(indices_p) %>% 
    summarize(mean_jitter = mean(pain)) %>% 
    select(mean_jitter) %>% 
    pull()

  boot_diffs[i] <- temp_h - temp_p
}
boot_diffs <- tibble(diffs = boot_diffs)
boot_diffs %>% 
  summarize(lower = quantile(diffs, 0.025),
            upper = quantile(diffs, 0.975))
obs_diff <- boot_diffs %>% 
  summarize(obs_diff = mean(diffs)) %>% 
  pull()
offset <- boot_diffs %>% 
  summarize(offset = 0 - mean(diffs)) %>% 
  pull()
null_dist <- boot_diffs %>% 
  mutate(centered_diffs = diffs + offset) %>% 
  select(centered_diffs)
null_dist %>% 
  mutate(extreme = ifelse(centered_diffs > abs(obs_diff), 1, 0)) %>% 
  summarize(p_val = mean(extreme))

```


Since our p-value of (0.0055) is less than the alpha level of 0.05, we decide to
reject the null hypothesis that the mean throat pain after 30 after a licorice 
solution is greater than or equal to the mean pain after 30 with the sugar 
placebo solution, in favor of the alternative hypothesis that the mean throat 
pain after 30 after a licorice gargle is less than to the mean pain after 30 with
the sugar gargle.

We ultimately conclude that there is enough evidence to suggest a licorice gargle 
decreases the mean throat pain after 30 after a licorice gargle


### Exercise 6

Since this is an experimental study if we assume that those who have a sore throat
will have sore throat pain, we can conclude that a licorice solution will decrease 
sore throat pain 30 minutes after. This is because in exercise 4 we tested and 
found a p value of 1.804e-06 which suggests that there is significance to 
indicate that a licorice solution will decrease the sore throat pain 30 minutes 
after. Thus we can say the licorice solution prior to surgery is effective in r
educing post-intubation sore throat. 

However if we say that this conclusion holds true for longer periods of time (90
minutes or 4 hours or at 1 am), we can not say that is it effective. That is 
because we did not test for that duration of time and thus using the 30 minute
data would be an extrapolation. 

### Exercise 7
TRUE


### Exercise 8

FALSE. Our p-value can not be evidence for, but rather it would only be evidence
that does not reject it can not support. In addition we can not determine the
strength our data has in supporting the null hypothesis. 




### Exercise 9


False - we do not know the power of the test thus we can not find the (1 - power)
in order to determine the probability of having a type two error. 




### Exercise 10

FALSE- we are only 95% confident in that if we were to repeatedly 
construct 95% confidence intervals for the difference in mean BMI 
in the same way from the original population, then we know that 95% 
of those intervals would truly contain the true population difference in means.


### Exercise 11

True - the variance is already in the confidence interval

### Exercise 12


TRUE
